R Notebook for MSDepression Project

### 2021-11-05 lab_notebook/lab_notebook.Rmd
lab notebook additions

       	 	antsApplyTransforms -e 3 -d 3 -i ${mimosa_path} -o ${mimosa_mni_hcp_path} -r ${mni_t1_target} -t ${ms2mni_wa
rp} -t ${affine_mat} -n MultiLabel

@@

>>>>

       	 	antsApplyTransforms -e 3 -d 3 -i ${mimosa_path} -o ${mimosa_mni_hcp_path} -r ${mni_t1_target} -t ${ms2mni_wa
rp} -t ${affine_mat} -n GenericLabel

@@

20211019
Still having issues with transform, running slow, realizing that I had multilable instead of generic label. Will try that. Still worried I will have trouble getting it to work.

Got it to run by putting a bsub inside the script!

20211026
- Now trying to get dsi studio to run on pmacs. 
- scp -r /Users/eballer/msdepression/templates/dti to /project/msdepression/templates

Made a list of the paths to all the mimosa masks
ls /project/msdepression/data/subj_directories/*/*//*/mimosa*mn* > mimosa_binary_masks_hcp_space_20211026_n2336

Then moved to melissa's data directory
cp mimosa_binary_masks_hcp_space_20211026_n2336 ../melissa_martin_files/csv/.

20211027
 - fixing dsi studio. Goal today is to figure out how to bsub it because it takes a long time. steps

```{bash}
ssh -Y eballer@scisub.pmacs.upenn.edu
cd /project/msdepression/scripts
xbash

 Also needed to generate a list of niftis (i.e. tracts is voxel space). Did this locally and scp'ed to templates/dti

```{bash}
#!/bin/bash

for lesion in ${lesion_paths}; do
    echo "Working on file" $lesion
    # make directory within parent directory for tracts
    #parent_dir=$(echo $lesion | perl -pe s'/(.*)\/.*/$1/g')
    mask_prefix=$(echo $lesion | perl -pe s'/(.*)\/(.*).nii.gz/$2/g')
    if [ ! -d $parent_dir/fiber_tracking_maps ] 
    then
        mkdir $parent_dir/fiber_tracking_maps
    fi

    fiber_bundle_type_list=(
        "association"
        "cerebellum"
        "cranial_nerve"
        "projection"
        "commissural")
    for fiber_bundle_type in "${fiber_bundle_type_list[@]}"; do
        echo "Fiber bundle type is " $fiber_bundle_type
        echo "making directory" $parent_dir/fiber_tracking_maps/$fiber_bundle_type
        #mkdir $parent_dir/fiber_tracking_maps/$fiber_bundle_type
        fascicles=$(ls ${fascicle_directory}/${fiber_bundle_type}/*.tt.gz)
        for fascicle in ${fascicles}; do
            echo "Fascicle is  " $fascicle
            fascicle_root_name=$(echo $fascicle | perl -pe s'/(.*)\/(.*).tt.gz/$2/g') 
          #  dsi_studio --action=ana --source=$template --tract=$fascicle --roa=$lesion --output=$parent_dir/fiber_tracking_maps/${fiber_bundle_type}/${fascicle_root_name}_${mask_prefix}_lesioned_ROA.tt.gz
          #  dsi_studio --action=ana --source=$template --tract=$fascicle --roi=$lesion --output=$parent_dir/fiber_tracking_maps/${fiber_bundle_type}/${fascicle_root_name}_${mask_prefix}_lesioned_ROI.tt.gz
           # dsi_studio --action=ana --source=$template --tract=$fascicle --output=${fascicle_directory}/${fiber_bundle_type}/${fascicle_root_name}_${mask_prefix}_full.tt.gz
            dsi_studio --action=ana --source=$template --tract=$fascicle --output=${fascicle_directory}/${fiber_bundle_type}/${fascicle_root_name}.nii.gz
        done
        # $path/dsi_studio --action=ana --source=/Users/eballer/BBL/msdepression/templates/dti/HCP1065.1mm.fib.gz --tract=/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/association/SLF2_R.tt.gz --roa=/Users/eballer/BBL/msdepression/templates/perfect_ms_subject/run-001/mimosa_binary_mask_0.25.nii.gz --output=/Users/eballer/BBL/msdepression/templates/roi_files_for_testing/SLF2R_lesioned_ROA.nii.gz
        # $path/dsi_studio --action=ana --source=/Users/eballer/BBL/msdepression/templates/dti/HCP1065.1mm.fib.gz --tract=/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/association/SLF2_R.tt.gz --roi=/Users/eballer/BBL/msdepression/templates/perfect_ms_subject/run-001/mimosa_binary_mask_0.25.nii.gz --output=/Users/eballer/BBL/msdepression/templates/roi_files_for_testing/SLF2R_lesioned_ROI.nii.gz
        # $path/dsi_studio --action=ana --source=/Users/eballer/BBL/msdepression/templates/dti/HCP1065.1mm.fib.gz --tract=/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/association/SLF2_R.tt.gz --output=/Users/eballer/BBL/msdepression/templates/roi_files_for_testing/SLF2R_full_ROI.nii.gz
    done
done
```
 20211028

 Big moves! Currently running these two scripts and they have been git'ed
```{bash dsi_studio_bash.sh}
#!/bin/bash

#### Welcome to the Region_making party #########
### Pre: Must have a file with full paths to the lesioned data 
### Post: Within each directory specified by the lesioned data, will have a directory that has all tractography (ROA, ROI, Full)
### Uses: For use in MS depression - take a subject's mimosa lesions and generate the fiber tracts (individual fascicles) that run through it
#dependencies: Using dsi studio from docker, sif created by Tim 10/26/2021
#export PATH=${PATH}:/Applications/dsi_studio.app/Contents/MacOS/
set -euf -o pipefail

export ITK_GLOBAL_DEFAULT_NUMBER_OF_THREADS=$LSB_DJOB_NUMPROC
num_cores=1

#set default paths 
template='/project/msdepression/templates/dti/HCP1065.1mm.fib.gz'
default='/project/msdepression/data/melissa_martin_files/csv/mimosa_binary_masks_hcp_space_20211026_n2336'
#default='/project/msdepression/data/melissa_martin_files/csv/erica_mini_mimosa_paths_n3'
fascicle_directory='/project/msdepression/templates/dti/HCP_YA1065_tractography/'

# loop through each mimosa lesion map; lesion paths contain full paths to mimosa files in hcp space
job_count=1
for lesion in ${lesion_paths}; do
	echo "working on ${lesion} ..."  
	bsub -J "job_${job_count}" -n ${num_cores} -o /project/msdepression/scripts/logfiles/out_roa_${job_count}.out /project/msdepression/scripts/indiv_mimosa_lesion_dsi_s
tudio_script.sh $lesion
   	((job_count+=1))
done

and this calls 
```{bash indiv_mimosa_lesion_dsi_studio_script.sh}

### The Inner Script... Doing all the dsi studio action for each individual mimosa brain ###

### Pre: path to an individual's mimosa lesion in HCP/MNI space
### Post: Filtered tracts; ROA and ROI voxel maps/niftis that can be used for data analysis
### Uses: In trying to find a way to speed everything along, it made the most sense to separate this piece of the script out. 
#          - it takes a subject's mimosa path, and makes a fiber_tracking_map directory as well as sub directories for each bundle type
#          - it then runs ROA/ROI filtration and makes output maps.

#set -euf -o pipefail
module load singularity
export PATH=${PATH}:/Applications/dsi_studio.app/Contents/MacOS/
export ITK_GLOBAL_DEFAULT_NUMBER_OF_THREADS=$LSB_DJOB_NUMPROC
num_cores=1

#set default directory/templates. This will change when we use local, personal dwi
template='/project/msdepression/templates/dti/HCP1065.1mm.fib.gz'
fascicle_directory='/project/msdepression/templates/dti/HCP_YA1065_tractography/'

if [ $# == 0 ]
then
    echo "You did not enter a file. Make sure your call makes sense"
else  
    lesion=$1

	echo "File being read is "$lesion
	echo "... Starting to make tracts"
# make directory within parent directory for tracts
    	parent_dir=$(echo ${lesion} | perl -pe s'/(.*)\/.*/$1/g')
    	echo "$parent_dir"
	mask_prefix=$(echo ${lesion} | perl -pe s'/(.*)\/(.*).nii.gz/$2/g')
    	echo $mask_prefix
	if [ ! -d $parent_dir/fiber_tracking_maps ] 
    	then
        	mkdir $parent_dir/fiber_tracking_maps
    	fi

    #make sub directories for each bundle type, so within fiber_tracking_maps, there will be subdirectories for each type of map
	fiber_bundle_type_list="association cerebellum cranial_nerve projection commissural"	
    	for fiber_bundle_type in ${fiber_bundle_type_list}; do
        	echo "Fiber bundle type is " $fiber_bundle_type
        	echo "making directory" $parent_dir/fiber_tracking_maps/$fiber_bundle_type
        	mkdir $parent_dir/fiber_tracking_maps/$fiber_bundle_type
        	fascicles=$(ls ${fascicle_directory}/${fiber_bundle_type}/*.tt.gz)

	# set a job counter to be used to name the jobs
		job_count=1
		for fascicle in ${fascicles}; do
            		echo "Fascicle is  " $fascicle
            		fascicle_root_name=$(echo $fascicle | perl -pe s'/(.*)\/(.*).tt.gz/$2/g') 
   			echo  "Fascicle root name is ${fascicle_root_name}" 
    	    		singularity exec --bind /project /project/singularity_images/dsistudio_latest.sif dsi_studio --action=ana --source=$template --tract=$fascicle --roa=
$lesion --output=$parent_dir/fiber_tracking_maps/${fiber_bundle_type}/${fascicle_root_name}_${mask_prefix}_lesioned_ROA.tt.gz
            		singularity exec --bind /project /project/singularity_images/dsistudio_latest.sif dsi_studio --action=ana --source=$template --tract=$fascicle --roi=
$lesion --output=$parent_dir/fiber_tracking_maps/${fiber_bundle_type}/${fascicle_root_name}_${mask_prefix}_lesioned_ROI.tt.gz

     			singularity exec --bind /project /project/singularity_images/dsistudio_latest.sif dsi_studio --action=ana --source=$template --tract=$fascicle --roa=
$lesion --output=$parent_dir/fiber_tracking_maps/${fiber_bundle_type}/${fascicle_root_name}_${mask_prefix}_lesioned_ROA.nii.gz
           		singularity exec --bind /project /project/singularity_images/dsistudio_latest.sif dsi_studio --action=ana --source=$template --tract=$fascicle --roi=
$lesion --output=$parent_dir/fiber_tracking_maps/${fiber_bundle_type}/${fascicle_root_name}_${mask_prefix}_lesioned_ROI.nii.gz

		# I have removed this line to save space. All hcp fibers are in the directory: /project/msdepression/templates/dti/HCP_YA1065_tractography/ . Can uncomment t
his if at some point I need to make them for each individual, like when we have their personal DWI
#singularity exec --bind /project /project/singularity_images/dsistudio_latest.sif dsi_studio --action=ana --source=$template --tract=$fascicle --output=$parent_dir/fiber_tr
acking_maps/${fiber_bundle_type}/${fascicle_root_name}_${mask_prefix}_full.tt.gz
			((job_count+=1))
		done
    	done
fi
```

20211029
[x] 3dmaskave, get the # of voxels into csvs

scripts are made, just have no more space on pmacs, so waiting for them to make space for me and then I can do this.

20211103
[x] start working on getting these files into r, MAKE SURE TO ADD A ZERO in the read in line = NA -> 0!!!

20211104
[x] ton of weird errors that are inconsistent with afni. Need to figure out if they are real or fake.

Here is how I figured out the failed subj/session #
```{bash}
more 3dmaskav_*.out | grep "ERROR" | perl -pe 's/.*\/project.*\/(sub-.*\/ses-.*)\/run.*/$1/' | sort -u > failed_3dmaskave_n243_20211103
```

Have identified the people who failed, it looks like their lesions are empty for these
```{bash}
more failed_3dmaskave_n243_20211103 | perl -pe 's/(.*)/\/project\/msdepression\/data\/subj_directories\/$1\/run-001\/mimosa_binary_mask_0.25_mni_hcp.nii.gz/' > failed_full_mimosa_paths_n243
```

[x]Going to try rerunning these specifically from the dsi_studio command, using these files instead of what is already there

Have to make a file that contains all the full volumes of each of the fascicles

```{bash make_streamline_volumes_for_template.sh}
### Make streamline volumes ###

### Pre: All template streamlines must have been made in dsi studio (these are the nii files)
### Post: fiber_volume_values.csv that contains the name of the fiber bundle and its volume
### Uses: Will be getting the volumes of the streamlines for use in later density analyses
### Dependencies: This script requires afni to be installed.

module load afni_openmp/20.1
export ITK_GLOBAL_DEFAULT_NUMBER_OF_THREADS=$LSB_DJOB_NUMPROC
num_cores=1

directory='/project/msdepression/templates/dti/HCP_YA1065_tractography/'
outfile=${directory}/fiber_volume_values.csv

#remove outfile if it exists
rm -f ${outfile}

#make the file again
touch ${outfile}
echo "fascicle,volume" >> ${outfile}

#go through each fiber, get volume, and store it in csv
fiber_bundle_type_list="association cerebellum cranial_nerve projection commissural"	
for fiber_bundle_type in ${fiber_bundle_type_list}; do
	echo "Fiber bundle type is " $fiber_bundle_type
fascicle_volumes=$(ls ${directory}/${fiber_bundle_type}/*.nii*)
        for fascicle in ${fascicle_volumes}; do
            fascicle_prefix=$(echo $fascicle | perl -pe 's/.*\/(.*).nii.*/$1/g')
            volume=$(3dmaskave -quiet -mask SELF -sum ${fascicle})
            echo "${fascicle_prefix},$volume" >> ${outfile}
        done
done
```

output of the above script
```{bash}
fascicle,volume
AF_L,35716
AF_R,15249
C_FPH_L,2569
C_FPH_R,2760
C_FP_L,14457
C_FP_R,21386
C_PH_L,3171
C_PHP_L,1900
C_PHP_R,4612
C_PH_R,2593
C_R_L,2666
C_R_R,5181
EMC_L,7344
EMC_R,5641
FAT_L,27142
FAT_R,15510
IFOF_L,30311
IFOF_R,34880
ILF_L,27401
ILF_R,36327
MdLF_L,7637
MdLF_R,5744
PAT_L,10921
PAT_R,13976
SLF1_L,9836
SLF1_R,13482
SLF2_L,26947
SLF2_R,37091
SLF3_L,10053
SLF3_R,15526
UF_L,7607
UF_R,7744
VOF_L,5681
VOF_R,4079
CB_L,45470
CB_R,43251
ICP_L,4963
ICP_R,4449
MCP,16514
SCP,4315
V,11299
CNIII_L,563
CNIII_R,486
CNII_L,1097
CNII_R,1415
CNVIII_L,1034
CNVIII_R,968
CNVII_L,298
CNVII_R,276
CNV_L,817
CNV_R,995
AR_L,3082
AR_R,1935
CBT_L,4609
CBT_R,6146
CPT_F_L,18693
CPT_F_R,34955
CPT_O_L,9022
CPT_O_R,7015
CPT_P_L,24377
CPT_P_R,27829
CS_A_L,26114
CS_A_R,21542
CS_P_L,5526
CS_P_R,6231
CS_S_L,31034
CS_S_R,34141
CST_L,20476
CST_R,16388
DRTT_L,14986
DRTT_R,17163
F_L,5175
F_R,8018
ML_L,10871
ML_R,9139
OR_L,6876
OR_R,5510
RST_L,14766
RST_R,11890
TR_A_L,20628
TR_A_R,14840
TR_P_L,4402
TR_P_R,4748
TR_S_L,39767
TR_S_R,37036
AC,26517
CC,390601
```

2021104
[x] need to see if total number of fibers in a bundle = ROA + ROI - NO
[] if all these work, our first analyses will be depressed vs non total , total fiber volume lost in lesions, then specificity. WOuld look at ratios ROI/total

To initialize Rstudio on pmacs
```{bash}
source  /project/bbl_projects/apps/default_modules.sh
conda activate rstudio
xbash
rstudio
### OR ###
cd /project/msdepression/scripts
source load_rstudio
```

To make list of all roi files
```{bash}
cd /project/msdepression/data/melissa_martin_files/csv/
more mimosa_binary_masks_hcp_space_20211026_n2336 | perl -pe 's/(.*)mimosa_bin.*/$1fiber_tracking_maps\/fiber_volume_values_roi\.csv/' > roi_fiber_volumes_n2336
```

Rscript to make df with lesions

```{r Roi_ratio_regressions}
### ROA Ratio Regressions ###

#####
## Pre: /project/msdepression/data/melissa_martin_files/csv/roi_fiber_volumes_n2336, which contains full paths to the ROI files, can be ammended to do ro
## Post: Analysis
## Uses: 1) reads in the ROIs, puts them in a data frame (columns are accession num/date/each fascicle), one row per subjected per date
#        2) Performers univariate analyses, depression diagnosis versus not
## Dependencies: R 3.6.3

#include packages
library(dbplyr)

#set directories
homedir<-'/project/msdepression/'

#analysis type
region_type="roi"

#read in template fascicle volumes
template_full_volumes <- read.csv(paste0(homedir, '/templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv'), header = T)
fascicle_names <- template_full_volumes$fascicle

### Read in file with all full paths to region volumes
region_volume_csv <- read.table(paste0(homedir, '/data/melissa_martin_files/csv/', region_type, '_fiber_volumes_n2336'), header = F, stringsAsFactors = F)
region_volume_csv$V1 <- as.character(region_volume_csv$V1)

#outfile, named for number of subjects
outfile <- paste0(homedir, "/results/fascicle_volumes_all_subjects_", region_type, "_n", dim(region_volume_csv)[1], ".csv")

#make data frame, 1 row per subject, num rows = num rows in region_volume_csv
volumes_df <- data.frame(matrix(nrow = dim(region_volume_csv)[1], ncol = 89))
names(volumes_df) <- c("ACCESSION_NUM", "EXAM_DATE", as.character(fascicle_names))
#loop through region_volume_csv, read in df corresponding to each row of region_volume_csv, extract accession num and date,
## and transpose fiber volumes from column to rows. append accession num/date/region_volumes row into big matrix

for (subj in 1:dim(region_volume_csv)[1]) {
  file_path <- region_volume_csv$V1[subj]
  region_volume_data <- read.csv(file_path, header = T, sep = ",")
  accession_num<- region_volume_data[1,1]
  date <- region_volume_data[1,2]
  volumes <- t(region_volume_data[,4])
  volumes[is.na(volumes)] <- 0
  volumes <- volumes/template_full_volumes$volume
  row_to_add <- cbind(accession_num, date, volumes)
  volumes_df[subj,]<- row_to_add
}

write.csv(file=outfile, volumes_df, quote = FALSE, row.names = FALSE)

To see the frequencies of the lesions
```{bash MC plot}
library(reshape2)
library(ggplot2)
df <- read.csv("fascicle_volumes_all_subjects_roi_n2336.csv", header=TRUE)

m.df <- melt(df, id.vars = c("ACCESSION_NUM", "EXAM_DATE"))
lesioned <- subset(m.df, value > 0)

ggplot(lesioned, aes(x=value)) + geom_histogram() + facet_wrap(~variable)
```

20211107
[] need to change the date format in my regular dac pull
[] need to divide the ROIs by the total volume 
[] need to get total # of lesions - this will be tough, need taki's code. Not to do right now - sydney to look into
